---
title: "Jellycast"
---

```{r setup, echo = FALSE}
source("setup.R")
coast = read_coast()
coast_bbox = st_bbox(coast)
obs = read_obs() %>% 
  filter(source == "record", type %in% c("lionsmane", "moon", "whitecross", "comb")) %>% 
  st_as_sf(coords = c("lon", "lat"), crs = 4326) %>% 
  mutate(year = year(date),
         month = factor(month(date, label = TRUE, abbr = TRUE), 
                        levels = month.abb),
         type = as.factor(type))
```

For the [gom-jellies](https://github.com/BigelowLab/gom-jellies) project, observations come from community scientist reports. The following species have been selected for analysis:

  + *Cyanea capillata* (Lion's Mane)
  + *Aurelia aurita* (Moon)
  + *Staurostoma mertensii* (White Cross)
  + *Ctenophores* (Comb)

# All Observations
```{r}
ggplot() +
  geom_sf(data = coast, color = "black") +
  geom_sf(data = obs, color = "red", size = 1.5) +
  coord_sf(
    xlim = c(coast_bbox["xmin"], coast_bbox["xmax"]),
    ylim = c(coast_bbox["ymin"], coast_bbox["ymax"]),
    expand = FALSE
  ) +
  theme_minimal() +
  labs(title = "All Observations") +
  theme(axis.title = element_blank())
```

# Observations by Year
```{r}
ggplot() +
  geom_sf(data = coast, color = "black") +
  geom_sf(data = obs, color = "red", size = 1.5, alpha = 0.7) +
  coord_sf(
    xlim = c(coast_bbox["xmin"], coast_bbox["xmax"]),
    ylim = c(coast_bbox["ymin"], coast_bbox["ymax"]),
    expand = FALSE
  ) +
  facet_wrap(~ year) +
  theme_minimal() +
  labs(title = "Observations by Year") +
  theme(axis.title = element_blank())
```

# Observations by Month
```{r}
ggplot() +
  geom_sf(data = coast, color = "black") +
  geom_sf(data = obs, color = "red", size = 1.5, alpha = 0.7) +
  coord_sf(
    xlim = c(coast_bbox["xmin"], coast_bbox["xmax"]),
    ylim = c(coast_bbox["ymin"], coast_bbox["ymax"]),
    expand = FALSE
  ) +
  facet_wrap(~ month) +
  theme_minimal() +
  labs(title = "Observations by Month") +
  theme(axis.title = element_blank())
```

# All Observations by Species
```{r}
ggplot() +
  geom_sf(data = coast, color = "black") +
  geom_sf(data = obs, aes(color = type), size = 1.5) +
  coord_sf(
    xlim = c(coast_bbox["xmin"], coast_bbox["xmax"]),
    ylim = c(coast_bbox["ymin"], coast_bbox["ymax"]),
    expand = FALSE
  ) +
  theme_minimal() +
  labs(title = "All Observations") +
  theme(axis.title = element_blank())

```

# Observations by Year
```{r}
ggplot() +
  geom_sf(data = coast, color = "black") +
  geom_sf(data = obs, aes(color = type), size = 1.5, alpha = 0.7) +
  coord_sf(
    xlim = c(coast_bbox["xmin"], coast_bbox["xmax"]),
    ylim = c(coast_bbox["ymin"], coast_bbox["ymax"]),
    expand = FALSE
  ) +
  facet_wrap(~ year) +
  theme_minimal() +
  labs(title = "Observations by Year") +
  theme(axis.title = element_blank())
```

# Observations by Month
```{r}
ggplot() +
  geom_sf(data = coast, color = "black") +
  geom_sf(data = obs, aes(color = type), size = 1.5, alpha = 0.7) +
  coord_sf(
    xlim = c(coast_bbox["xmin"], coast_bbox["xmax"]),
    ylim = c(coast_bbox["ymin"], coast_bbox["ymax"]),
    expand = FALSE
  ) +
  facet_wrap(~ month) +
  theme_minimal() +
  labs(title = "Observations by Month") +
  theme(axis.title = element_blank())
```